---
title: "Results in plots"
subtitle: "Plot SWAT+ model calibration results"
output: html_document
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The calibration of the SWAT+ model is a complex process that involves many steps. One of the most important steps is the visual inspection of the model results. This document provides an overview of the functions available in the `SWATtunR` package for plotting SWAT+ model calibration and validation results. Functions intended to be used for plotting at different stages of the SWAT+ model calibration and validation process and not in sequential order. To run these function [step 6](#hc_step7) or/and [step 6a](#hc_step6a) should be completed successfully. Same packages should be loaded as in the [step 1](#hc_step1).

## 1. Dotty plots {#hcp_step1}

For comparison, you can use the function [plot_dotty](../reference/plot_dotty.html). It allows examining model performance results versus parameter values.

The function has the following arguments:

- `par`: a data frame of model parameter values for each simulation.
- `var`: model performance result vector or list of vectors to be plotted against parameter values.
- `y_label`: the label or vector of labels for the y-axis. Default is `'y'`.
- `n_col`: number of columns in the facet grid. Default is `3`.
- `y_lim`: defines limits for the y-axis. Default is `NULL`.
- `y_inter`: Y-axis intercept value. Default is `NULL`.
- `trend`: logical, indicating whether to add a trend line. Default is `FALSE`.
- `run_ids`: a numeric vector of run IDs to be highlighted in the plot.
- `low_up`: logical, TRUE if the whole possible parameter range should be used for the x-axis. Default is `FALSE`.

```{r dotty, message = FALSE, warning = FALSE, eval = FALSE}
# Example of usage:
plot_dotty(sim_flow$parameter$values, obj_tbl$nse)
plot_dotty(sim_flow$parameter$values, obj_tbl$kge, n_col = 5, y_lim = c(0.5, 0.9))
plot_dotty(sim_flow$parameter$values, obj_tbl$pbias, n_col = 5)
plot_dotty(sim_flow$parameter$values, obj_tbl$nse, run_ids = c(1, 2, 5), low_up = TRUE)
plot_dotty(sim_flow$parameter$values, obj_tbl$kge, trend = TRUE)
```

Dotty plots can also be created for multiple variables in the simulation list. In this case, the dotty plot will be displayed on the same plot with different colors for each variable. Trends can also be added to the plot. Water flow and water quality variables can be plotted together on the same plot.

```{r dotty_multi, message = FALSE, warning = FALSE, eval = FALSE}
# Example of usage:
plot_dotty(par = sim_flow$parameter$values, 
           var = list(obj_tbl_list$flo_day_52$nse, 
                      obj_tbl_list$no3_day_52_conc$nse,
                      obj_tbl_list$gwd_day$nse), 
           y_label =  c('Flow at 52r (m3/s)', "Nitrate at 52r (mgN/l)", "Groundwater depth (m)"),
           trend = TRUE)
```

## 2. Simulation vs observed values time series {#hcp_step2}

To plot the results of selected variable sets simulation results versus observed, you can use the [plot_selected_sim](../reference/plot_selected_sim.html). This function allows you to compare the selected simulated data with the observation data. The function has several arguments:

- `sim`: object from SWATrunR.
- `obs`: dataframe for the observed data with two columns: 'date' and 'value'.
- `par_name`: name of the parameter set to be used (default is NULL, which selects the first record). If you have multiple parameter sets, you can use this argument to select the one you want to use. For example, `par_name = "flo_day"` or `par_name = "no3_day_conc"`.
- `run_ids`: character vector with the IDs of the runs to be used in the plot.
- `run_sel`: character vector with the ID or IDs of the runs to be emphasized in the plot.
- `plot_bands`: logical. If TRUE, the plot will include lower and upper bands of results.
- `period`: character. If NULL, the plot will be based on the original time step of the simulation. Other examples are "day", "week", "month", "year", etc. If "average monthly", the plot will be based on the multi-annual monthly values.
- `fn_summarize`: function to recalculate the time interval (default is "mean"), other options include "mean", "median", "sum", "min", "max", etc.
- `x_label`: character. Label for the x-axis. Default is "Date".
- `y_label`: character. Label for the y-axis. Default is "Discharge (m³/s)".

```{r selected_sim, message = FALSE, warning = FALSE, eval = FALSE}
##Simple example
plot_selected_sim(sim_flow, obs, par_name = "flo_day_52", run_ids = c(1, 2, 3))

# Another example how all function arguments are used
plot_selected_sim(sim = sim_flow,
                  obs = obs,
                  par_name = "flo_day",
                  run_ids = run_sel_ids,
                  run_sel = run_sel_ids[1],
                  plot_bands = TRUE,
                  period = "average monthly",
                  fn_summarize = "mean",
                  x_label = "Month",
                  y_label = "Monthly flow average (m<sup>3</sup> s<sup>-1</sup>)")
```

## 3. Plotting parameter identifiability {#hcp_step3}

The plot below shows the ranges of parameter,s which perform better for a certain performance metric. The `run_fraction` defines the fraction of simulations which should be considered in the analysis (with 1000 runs this value should maybe not be lower than 0.2, too large values also do not make much sense). For 5000 runs I think 0.1 would be a good value (I think this is what Bjoern used in his publication).

In case you used multiple variables in performance calculation, you can use `obj_tbl_list$var_name` to select the variable for which you want to plot parameter identifiability. Similar adjustment could be done for dotty plots.

```{r identifiability, message = FALSE, warning = FALSE, eval = FALSE}
# Usage example
plot_parameter_identifiability(parameters = sim_flow$parameter$values,
                               objectives = 
                                 select(obj_tbl, -starts_with(c('rank', 'run'))), 
                               ## This starts_with vector could be amended with
                               ## other columns which should not be considered
                               run_fraction = .2)

```

