---
title: "Performing Runs"
subtitle: "Running SWAT simulations for hard calibration"
output: html_document
bibliography: ref.bib
---

This page corresponds to **03_run_swat.R** script in the hard calibration workflow. It runs the SWAT+ model using the defined outputs and parameters.

# 4. Perform SWAT simulation runs

Firstly, following settings could be modified based on the user's requirements.

```{r hc4, message = FALSE, warning = FALSE, eval = FALSE}
# Parameter definition ----------------------------------------------------
# Path to the SWAT+ project folder.
model_path <- 'test/swatplus_rev60_demo'

# Start date of simulation period
start_date <- '2003-01-01'
# End date of simulation period
end_date <- '2012-12-31'
# Start date for printing simulation outputs
start_date_print <- '2006-01-01'

# Number of cores used for parallel simulation runs
n_cores <- NULL

# Name of the folder where simulation results will be saved incrementally.
# To continue writing to existing saved runs, replace by the name of the
# existing save_file.
save_file_name <- paste0(format(Sys.time(), '%Y%m%d%H%M'), '_sim')

# Path where the simulation results are saved.
# Default the simulations are saved in the calibration project
# in the sub-folder /simulation
save_path <- './simulation'
```

The following section runs the SWAT+ model using the defined outputs and parameters. The `run_swatplus` function from **SWATrunR** package is used to execute the model runs. 

```{r hc41, message = FALSE, warning = FALSE, eval = FALSE}
# Perform simulation runs
run_swatplus(project_path     = model_path,
             output           = outputs,
             parameter        = parameter_set,
             start_date       = start_date,
             end_date         = end_date,
             start_date_print = start_date_print,
             n_thread         = n_cores,
             save_path        = save_path,
             save_file        = save_file_name,
             return_output    = FALSE,
             split_units      = FALSE, # better set TRUE for large number of units
             time_out         = 3600 # seconds, change if run-time differs
             )
```
