---
title: "Hard calibration"
subtitle: "Hard calibration workflow for SWAT+ models"
output: html_document
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The calibration of the SWAT+ model is crucial for ensuring its accuracy and reliability in simulating hydrological processes and predicting the impacts of land management practices on water resources. Calibration involves adjusting the model's parameters to match observed data, such as streamflow, sediment yield, and nutrient loads, thereby improving its representation of the real-world watershed conditions. This process enhances the model's predictive capability, making it a more effective tool for water resource management, policy-making, planning and in the context of assessing the effects of climate change, land use alterations, and conservation practices. This page provides steps and tools to perform a hard calibration of SWAT+ models using scripted workflows.

## 1. Loading required packages {#hc_step1}

The first step in the hard calibration workflow is to load the necessary R packages. These packages provide functions and tools for data manipulation, visualization, and model calibration. The following code chunk loads the required packages for the hard calibration workflow.

```{r libraries_hc, message = FALSE, warning = FALSE}
library(data.table)
library(ggplot2)
library(hydroGOF)
library(lhs)
library(lubridate)
library(purrr)
library(readr)
library(stringr)
library(SWATrunR)
library(tibble)
library(tidyr)
library(parallel)
library(doParallel)
library(scales)
library(xts)
library(dygraphs)
library(SWATprepR)
library(dplyr)
```

## 2. Defining settings {#hc_step2}

The next step is to define the settings for the hard calibration workflow. This includes specifying the SWAT+ model path, input files, calibration parameters, observed data, and other relevant information. The following code chunk sets up the necessary settings for the hard calibration workflow.

```{r settings_hc, message = FALSE, warning = FALSE}
# Path to the SWAT+ model
model_path <- '../test/my_dearest_model'

# Path to the crop data file
qobs_path <- '../inst/extdata/q_cha.csv'

# Define the path to the observed WQ data
wqobs_path <- '../inst/extdata/no3.csv'

# You can define the path to the other observed data, like groundwater depth data, etc.
gwobs_path <- '../inst/extdata/gw.csv'

# Set the path to save results of the soft calibration
sc_res <- '../test/simulations'

# Set the number of cores available for calculations.  
# Number of cores for 12 runs use 3, 4, 6, or 12 cores (if available)
cores <- 3

# Channel IDs where gauges are located at the channel outlets.
cha_ids <- c(67)

# Define the simulation save file names to be used in SWATrunR
save_file_name <- "sim1"

# Number parameter combinations to run
n_comb <- 500

# Fraction of the best parameter combinations to be used for the final 
# WQ calibration
n_best <- NULL # example n_best <- .02 # 2% of the best parameter combinations, 
# if NULL, n_best = 20/n_comb

# Number of combination drawn for the nutrient parameters (this number 
# is multiplied by the number of selected parameter combinations for the flow.
# For example n_comb = 1000, n_best .1, means 100 best runs selected, multiplied
# with n_nutr = 100 results in 10,000 model runs for the nutrient calibration.
# If null, 1000/(n_comb*n_best) formula applied resulting in 1000 runs)
n_nutr <- NULL

# Number of cores to be used for the parallelization of the model runs
n_cores <- NULL ## If NULL, all available physical cores - 2 will be used

# Define time period to be used in model runs. Please note this period should
# match your management data period, if it was prepared with SWATfarmR. You can
# safely modify 'start_date_print' and 'end_date' to define the period for which
# you want to print the model output. However, please take great care, if you 
# modify 'start_date' as you might need to update your management files too. 

# Start date for model runs (if NULL, the start date of the model setup files 
# will be used!!!)
start_date <- NULL # start_date <- '2000-01-01'
# End date for model runs (if NULL, the end date of observation data will be used).
end_date <- NULL # end_date <- '2010-12-31'
# Start printing model output (if NULL, the start date of observation data will used).
start_date_print <- NULL # start_date_print <-'2003-01-01'

# Define calibration and validation periods. Please note that the validation 
# should take around 1/3 and calibration 2/3 of the total period. Also the 
# workflow will run all simulations for the whole period (defined above), 
# these periods are only used for the performance calculation and plotting.

# Q calibration period
q_cal_period <- c('2007-01-01', '2010-01-01')
# Calibration end date
q_val_period <- c('2011-01-01', '2022-12-31')

# WQ calibration period
wq_cal_period <- c('2012-01-01', '2017-12-31')
# Calibration end date
wq_val_period <- c('2018-12-31', '2022-12-31')
```

## 3. Preparing parameters sets for the flow (and water quality) calibration {#hc_step3}

Adding text here


```{r sampling, message = FALSE, warning = FALSE, eval = FALSE}
par_bound <- tibble(
  # Suggested list and ranges of SWAT+ paramaters for hydrology calibration in 
  # OPTAIN. Parameters are roughly divided into groups by processes which they 
  # affect most. Some groups should be considered optional
  # snow (optional - use if average snow fall to precipitation ratio is higher 
  # than 5%)
  'snomelt_tmp.hru | change = absval' = c(-1.5, 1.5),
  'snofall_tmp.hru | change = absval' = c(-1.5, 1.5),
  # ET (note: it is suggested that a narrow range for esco selected in soft 
  # calibration of water balance is used instead of the wide (0,1) range)
  'esco.hru | change = absval' = c(0.01, 0.95),
  'epco.hru | change = absval' = c(0.05, 1),
  'awc.sol | change = relchg' = c(-0.25, 0.25),
  'canmx.hru | change = relchg' = c(-0.5, 0.5),
  #surface runoff
  'cn2.hru | change = relchg' = c(-0.15, 0.15),
  'cn3_swf.hru | change = absval' = c(0, 1),
  'ovn.hru | change  = relchg ' = c(-0.25, 0.25),
  'surlag.bsn | change = absval' = c(0.05, 4),
  # lateral flow (optional - use if lateral flow constitutes at least 5% of total 
  # water yield)
  #'lat_time.hru | change = relchg' = c(-0,15, 0.15),
  'lat_len.hru | change = abschg' = c(-30, 30),
  'latq_co.hru | change = absval' = c(0, 1),
  'bd.sol | change = relchg' = c(-0.25, 0.25),
  'k.sol | change = relchg' = c(-0.5, 2),
  # tile flow (optional - use if tile flow constitutes at least 5% of total water
  # yield; note: tile_lag and tile_dtime should be active only if tile_drain is 
  # set to 0 in codes.bsn file))
  #'tile_dep.hru | change = relchg' = c(-0.1, 0.2),
  #'tile_lag.hru | change = absval' = c(48, 100),
  #'tile_dtime.hru | change = absval' = c(48, 100),
  #percolation/aquifer
  'perco.hru | change = absval' = c(0, 1),
  'flo_min.aqu | change = abschg' = c(-2, 2),
  'revap_co.aqu | change = absval' = c(0.02, 0.2),
  'revap_min.aqu | change = abschg' = c(-2, 2),
  'alpha.aqu | change = absval' = c(0.001, 0.5),
  # 'sp_yld.aqu | change = absval' = c(0.001, 0.05),
  'bf_max.aqu | change = absval' = c(0.5, 2),
  #channel routing
  'chn.rte | change = absval' = c(0.02, 0.1)
)

# Draw a first parameter set of parameter combinations 
par_flow <- sample_lhs(par_bound, n_comb)
```

## 3a. Preparing parameters sets for the flow (and water quality) calibration {#hc_step3a}

### Initialization of Parameters
The parameters `cn3_swf` and `latq_co` are initialized based on the runoff potential of a Hydrologic Response Unit (HRU), while the parameter `perco` is initialized based on its leaching potential. The calibration of these three parameters requires considering their different initial values. Therefore, the parameter sampling will be conducted as follows:

### Base Parameter Set

- For the parameters `perco`, `cn3_swf`, and `latq_co`, normalized ranges (*0*, *1*) are sampled.
- All other parameters are sampled within their actual parameter ranges.
- A base Latin Hypercube Sampling (LHS) set is drawn for the defined parameter boundaries.

### Translating Base Parameter Values
- From hydrology.hyd, HRU IDs with low, moderate, and high leaching and runoff potentials are drawn.
= These IDs are compiled into text strings required for parameter definition via the parameter names.
- Parameter boundaries for the different low, moderate, and high potentials are defined.
- The normalized parameter ranges are translated into the defined boundaries.

### Base Parameter Boundary Definition
This is a very reduced set for testing purposes and does not include all parameters that should be used in OPTAIN model setups.

*Note*: Calibration and validation of all hydrology and water quality/load related parameters could be done in one step. For this, you need to define the nutrient parameters together with hydrology. Please refer to step 12 for an example of how to define sediment, nitrogen, and phosphorus parameters. The example here presents only hydrology parameters.

```{r sampling_a, message = FALSE, warning = FALSE, eval = FALSE}
# Read the hydrology.hyd file to identify the HRU IDs with the different 
# runoff and leaching potentials.
hyd_hyd <- read_tbl(paste0(model_path, '/hydrology.hyd'))

# Generate the ID text strings to be included in the parameter names so that 
# parameter values for HRUs can be addressed separately.
init_perco <- c(low = 0.01, mod = 0.50, high = 0.95)

id_lch_pot <- c(low  = group_values(which(hyd_hyd$perco == init_perco['low'])),
                mod  = group_values(which(hyd_hyd$perco == init_perco['mod'])),
                high = group_values(which(hyd_hyd$perco == init_perco['high'])))

init_cn3 <- c(low = 0.95, mod = 0.30, high = 0.00)

id_run_pot <- c(low  = group_values(which(hyd_hyd$cn3_swf == init_cn3['low'])),
                mod  = group_values(which(hyd_hyd$cn3_swf == init_cn3['mod'])),
                high = group_values(which(hyd_hyd$cn3_swf == init_cn3['high'])))

# Define parameter boundaries in which the parameter values should vary for 
# the different leaching and runoff potentials.
perco_bound <- list(low = c(0.05, 0.30), mod = c(0.30, 0.60), high = c(0.5, 0.90))
cn3_bound   <- list(low = c(0.50, 0.95), mod = c(0.15, 0.45), high = c(0.0, 0.30))
latq_bound  <- list(low = c(0.01, 0.30), mod = c(0.10, 0.40), high = c(0.5, 0.90))

# Translate the normalized parameter ranges into the defined boundary ranges.
for(i in c('low', 'mod', 'high')) {
  if(nchar(id_lch_pot[i]) > 0) {
    par_flow[[paste0('perco_', 
                     i, '::perco.hru | change = absval | unit = c(', 
                     id_lch_pot[i], ')')]] <-  
      par_flow[['perco.hru | change = absval']] * (perco_bound[[i]][2] - perco_bound[[i]][1]) + perco_bound[[i]][1]
  }
  if(nchar(id_run_pot[i]) > 0) {
    par_flow[[paste0('cn3_', i, '::cn3_swf.hru | change = absval | unit = c(', 
                     id_run_pot[i], ')')]] <- 
      par_flow[['cn3_swf.hru | change = absval']] * (cn3_bound[[i]][2] - cn3_bound[[i]][1]) + cn3_bound[[i]][1]
    par_flow[[paste0('latq_', i, '::latq_co.hru | change = absval | unit = c(', 
                     id_run_pot[i], ')')]] <- par_flow[['latq_co.hru | change = absval']] * 
      (latq_bound[[i]][2] - latq_bound[[i]][1]) + latq_bound[[i]][1]
  }
}

# Remove the initial normalized parameters.
par_flow <- par_flow %>% 
  select(- starts_with(c('perco.hru', 'cn3_swf.hru', 'latq_co.hru')))
```
