---
title: "Soft Calibration Workflow "
output: html_document
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Loading required packages

The SWATtunR package is essential for soft calibration, as it provides the necessary functions for the calibration process. Additional packages are required for data manipulation, visualization, SWAT+ model runs, etc.

```{r load_packages, message = FALSE, warning = FALSE}
library(SWATtunR)
library(SWATrunR)
library(tidyverse)
```

## 2. Defining settings 

This step requires defining the SWAT+ model path, the path to the crop data file, and the observed water yield to precipitation ratio. 

For the crop yield soft calibration, it's essential to have accurate data on observed crop yields in dry weight. The data is typically sourced from a `'*.csv'` file, with columns specifying the plant names and their corresponding mean yields. If additional information is accessible, such as the minimum and maximum ranges for average annual observed yields, it's advisable to include them in the `'*.csv'` file as well, under the columns labeled yield_min and yield_max respectively. This ensures a comprehensive dataset for effective calibration and analysis.

```{r set_paths, message = FALSE, warning = FALSE}
# Path to the SWAT+ model
model_path <- '../test/my_dear_model'

# Path to the crop data file
crop_data_path <- '../inst/extdata/crop.csv'

# Observed water yield to precipitation ratio
obs_wy_ratio <- 0.27

# Set the path to save results of the soft calibration
sc_res <- '../test/simulations'
```

Crops observed yield data example is provided below.

```{r crop_data, message = FALSE, warning = FALSE}
##
crops_obs <- read_csv(crop_data_path, show_col_types = FALSE) %>% 
  filter(!is.na(yield_mean))

head(crops_obs, 3)
```

## 3. Days to maturity parameter changes

After settings following step allows to define, which days to maturity changes 
will be applied in soft-calibration. 

### 3.1. Initialization

Preparation of this step requires the conversion of observed yields information 
into numeric vectors and initialization of `'plants.plt'` file.

```{r init_dmat, message = FALSE, warning = FALSE}
# Convert observed yields to numeric vectors
for(i in c("yield_mean", "yield_min", "yield_max")){
  if(i %in% colnames(crops_obs)){
    assign(i, deframe(select(crops_obs, all_of(c("plant_name", i))))) 
  } else {
    assign(i, NULL)
  }
}

## Initialize unmodified backup version of plants.plt
if(!file.exists(paste0(model_path, '/plants.plt.bkp0'))) {
  copy_file_version(model_path, 'plants.plt', file_version = 0)
}
```

### 3.2. Days to maturity changes

The following code generates a parameter input table with changes for required 
for SWAT+ model run. The names are converted into [SWATrunR  syntax](https://chrisschuerz.github.io/SWATrunR/articles/par_sampl_calib.html) 
for the days to maturity parameter. The code snippet also samples changes for 
days to maturity, allowing users to define the crops to be used in soft calibration 
and the initial values for days to maturity. 

```{r dmat, message = FALSE, warning = FALSE}
# Reading the plants.plt input file from the initial unchanged backup file
plants_plt <- read_tbl(file_path = paste0(model_path, '/plants.plt.bkp0'))
plants_plt <- select(plants_plt, - any_of('description'))

# Define all crops which should be calibrated
# Alternatively define a name vector (if you also want to consider crops 
# which are not given in the observation input file).
crops <- crops_obs$plant_name
# crops <- c('crop1', 'crop2', 'crop3')

# Get a vector with days_mat initial values for the selected crops
dmat_init <- plants_plt %>%
  filter(name %in% crops) %>% 
  select(name, days_mat) %>%
  mutate(days_mat = ifelse(days_mat == 0, 110, days_mat)) %>% 
  ## if days_mat is 0, set it to 110 (as it is the default value)
  deframe()

# Sample changes for days to maturity. Depending on the number of cores you have 
# available choose a reasonable interval for the days to maturity which you want 
# to test. With an interval of 10 e.g. you have to do 12 simulations (3 rounds 
# on 4 cores), with an interval of 5 its 23 simulations.
dmat_step <- 10
dmat_chg <- round(seq(-30, 100, dmat_step))

# Generate the parameter input table with the changes for all crops and convert 
# the names into SWATrunR syntax
par_dmat <- map_df(dmat_init, ~ .x + dmat_chg) %>% 
  set_names(., paste0('dmat_', names(.),
                      '::days_mat.pdb | change = absval | name = ' , names(.)))

```

## 4. Run simulations

Days to maturity parameter changes are applied to the SWAT+ model, and simulations 
are run to analyze the PHU fractions, yields and biomass for the selected crops.
[SWATrunR](https://chrisschuerz.github.io/SWATrunR/) package is used to run the
simulations.

```{r run_simulations, message = FALSE, warning = FALSE, eval = FALSE, echo = FALSE}
# Set the number of cores. E.g. for 12 runs use 4 6, or 12 cores (if available)
cores <- 2

ylds_phu_dmat <- run_swatplus(project_path = model_path,
                              output = list(yld = define_output(file = 'mgtout',
                                                                variable = 'yld',
                                                                name = crops),
                                            bms = define_output(file = 'mgtout',
                                                                variable = 'bioms',
                                                                name = crops),
                                            phu = define_output(file = 'mgtout',
                                                                variable = 'phu',
                                                                name = crops)
                              ),
                              parameter = par_dmat,
                              start_date = '2015-01-01',
                              end_date = '2020-12-31',
                              years_skip = 2,
                              n_thread = cores,
                              time_out = Inf)

# Optionally save the simulation runs.
if(!dir.exists(sc_res)) dir.create(sc_res)
saveRDS(ylds_phu_dmat, paste0(sc_res, "/ylds_phu_mat.rds"))
```
